---
title: "Prediction Assignment Writeup"
output: html_document
---

## Background
Personal acelerometers have become increasingly popular for tracking activity levels for individuals with a wide variety of athletic background. In perticular the article "Qualitative Activity Recognition of Weight Lifting Exercises" (Velloso et al.) utilized a variety of accelerometers to quantify how well perticapents were able to complete excercises. Using the data collected for this study I was able to create a model to accuratly predict how an exercise was performed based on reedings from personal accelerometers.

### The Data
The data I used is available from http://groupware.les.inf.puc-rio.br/har along with a more detailed description of the study. After downloading the data and creating a training partition, I pull up the summary of the data frame so that I can begin to get an understanding of the data.

```{r, }
library(caret)
datafile<-read.csv("pml-training.csv")
set.seed(117)
inTrain <- createDataPartition(y=datafile$classe, p=0.60, list=FALSE)
training <- datafile[inTrain,]
testing <- datafile[-inTrain,]
example1<-summary(training$var_pitch_belt)
example2<-summary(training$avg_yaw_belt)
example1
example2
```
In the summary I noticed that several of the columns included empty or NA values. In particular I noticed that for each instance there were 11521 missing or NA values which indicated that perhaps that these values were all omited in the same rows so I decided to eliminate any column containing an omited value.
```{r}
key1<-is.na(datafile[1,])
training1<- training[,!key1]
key2<- training1[1,] == ""
training2 <- training1[,!key2]
dim(training2)
```
The "keys" were kept to ensure the same modifications can be made to the test sets. Finaly, the first 7 columns of the lates data frame consist of subject identifiers and timestamps:
```{r, echo=FALSE}
head(training2[,1:7])
```
Which are independent of the way the excercise is completed since each participant completed each excercise each way. THerefore, I elimintated these columns from the data.
```{r}
key3<- 8:60
training3<-training2[,key3]
dim(training3)
```
Finanly, I use this cleaned data set to compile a model to predict the classe based on the other variables.
```{r, cache=TRUE}
modelfit <- train(classe ~.,data=training3)
```

### Results
Looking at our model:
```{r}
modelfit
```
we see that we achieved 98.71% accuracy within are training set. By formatting and testing the testing set we have
```{r}
 testing1<-testing[,!key1]
testing2<-testing1[,!key2]
testing3<-testing2[,key3]
serror<-predict(modelfit, testing3)
confusionMatrix(serror, testing$classe)
```
And so w should be able to expect a 99% accuracy out of sample.